A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-03-28 CC/writeup.Rnw
more small changes.

\maketitle
\begin{abstract}
Abstract will be here
\end{abstract}
<<chunk_setup, echo=FALSE>>=
library(knitr)
opts_chunk$set(tidy = FALSE, echo=FALSE, cache=TRUE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
@
<<readingdata, eval=TRUE, echo=FALSE>>=
 library(ggplot2) # needs to be the version with geom_net!
 theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
 data(blood, football, theme_elements, lesmis)
 load("data/madmen.rda") # relative path should work for both of us
names(madmen$vertices)[1] <- "label"
names(blood$vertices)[1] <- "label"
names(theme_elements$vertices)[1] <- "label"
football$vertices$label <- as.factor(football$vertices$label)
football$vertices$value <- as.factor(football$vertices$value)
 email.net <- read.csv("data/email-data.csv")
 email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))

@@

This network is shown in figure~\ref{fig.cap:madmen_ex}. \emph{Notes: add names, color for gender}
<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men',cache=TRUE>>=
ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
  expand_limits(x = c(0,1.5), y = c(0,1)) + theme_net
@

@@

 In the plot, we can see one central character who has many more relationships than any other character. This vertex represents the main character of the show, Don Draper.
\par There are many kinds of networks, and networks are extensively studied across many disciplines.  Many sociologists study social networks, and many biologists study protein networks. As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.  We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}. \sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times: approximately 3454 downloads per day. It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.}

@@

  \subsection{Vertex Aesthetics}
  In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.  The \texttt{vlabel} aesthetic is the name or number assigned to each vertex in order to uniquely identify it.  This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be readable by the human eye.  The \texttt{vcolor} aesthetic can be an identity color (e.g. \texttt{color = "red"} outside of the \texttt{aes} statement) to change the color of all vertices, or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.  It can also color the vertices according to the different values of  some numeric variable, such as degree.  The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.
  %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs.  \footnote{Sam's questions/notes: Is ``degree" going to be a column in our data frame? Or will we have to write something to detect degree size?} It can also be set to any other numeric variable associated with the vertices in the network.  The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.  It can also change the shape of the vertices according to some grouping variable.  Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10, or to some numeric variable of interest that takes on the values $[0,1]$.

\par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion.
  \subsection{Edge Aesthetics}
  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for lines.  The \texttt{elabel} aesthetic is the name or number assigned to each edge.  This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.  The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to a grouping variable.  It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge. The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the many different types of dotted lines in \texttt{ggplot2}.  The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.  The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability.  Finally, the \texttt{directed} aesthetic is a logical value that identifies whether or not the network is directed.  When this value is true, the arrows change size, color, etc. with the previous aesthetics.

@@

\section{Many Many Examples}
\sctq{Format-wise, is it ok if each example is its own subsection? Also, can you check the references I added to the .bib file to make sure I'm creating them correctly? I've just been appending them.}\\
\sct{
In this section, we demonstrate the capabilities of \texttt{geom\_net} in a series of examples.
  \subsection{Blood Donation}
  \emph{Note: would be best with labels and direction}\\
  In this directed network, there are eight vertices and 27 edges.  The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive and negative types. \sctq{Can I cite wikipedia here? Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a persone whose blood type is that of a corresponding \emph{to} vertex. In the example below, loops are removed.  Loops exist on every vertex in this example, as blood of matching ABO and RhD type can always be received.
}
<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2)) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add direction later
@
\sct{
This network is shown in figure \ref{fig.cap:blood_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
\end{figure}

\subsection{Email Network}
  \emph{Note: would be best with coloring and direction}
  This email network comes from the 2014 VAST Challenge \citep{emailnet}. It is a directed network of emails between company employees with 55 vertices and 9,063 edges. Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees. The arrow of the directed edge points to the recipient(s) of the email. The network contains two business weeks of emails across the entire company. }
<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
ggplot(data = email.net, aes(from_id = From, to_id = to)) +
  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  theme_net #add direction later
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
\end{figure}

\sct{This network is plotted in figure \ref{fig.cap:email_ex}.  This plot is very crowded, and it is difficult to distinguish structure in the network when it there is so much over-plotting.  To remedy this, we facet the network by day to examine how the company's communications change throughout the week. }
<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  facet_wrap(~day, nrow = 2) +
  theme_net #add direction later
@
\begin{landscape}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
\caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
\end{figure}
\end{landscape}
\sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.  With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving? will know with direction option added later} several emails each day to a majority of the company. We can also see clusters that have formed on the perimeter of each day's network where the employees are communicating heavily with each other, but not much outside of their cluster. We can also see that there is heavier inter-cluster communication on different days.  For example, day 17 has more concentrated in-group communication than day 15. Additionally, we can better distinguish the communication of the outlying employee, who only corresponds with up to four other complany employees on a given day.} \sctq{talk about clusters being different departments within the company?}
  \subsection{\texttt{ggplot2} Theme Elements}
  \emph{Note: would be best with labels and direction} \\
  \sct{This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.  It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.  There are 53 vertices and 36 edges in this network.  There is an arrow from one vertex to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.  For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.  Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background. Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.  }
<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
  geom_net(vertices = theme_elements$vertices, vlabel = TRUE) +
  expand_limits(x = c(0,1), y = c(0,1)) +
  theme_net #add vlabel & direction later
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
\caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
\end{figure}

\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
  \subsection{Madmen Network (again?? or no?)}
  \sctq{leaving blank for now.}
  \subsection{College Football}
  \emph{Note: would be best with coloring vertices by conference}
  This next example comes from M.E.J. Newman's network data web page \citep{football}.  It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.  There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
ggplot(data = football$edges, aes(from_id = from, to_id = to, elinetype = in.conf + 1)) +
  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 2, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
  expand_limits(x = c(0,1), y = c(0,1)) +
  theme_net
@
\sct{
The network of football games is given in figure \ref{fig.cap:football_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
\end{figure}
  \subsection{Les Mis\'{e}rables}
  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
  This next network comes from \citet{lesmis}. It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.  There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.  There are 254 edges in this network. The edges are also weighted by the number of coappearances. The largest weighting is 31, between the characters Jean Valjean and Cosette.}
<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
\end{figure}
\sct{
  \subsection{Protein Interaction Network in Yeast}
  This example of a protein interaction network comes from \citet{protein}. It is the complete protein-protein interaction network in the yeast species \emph{S. cerevisiae}. There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.  These edges represent ``direct physical interactions" between any 2 proteins \citep[][p. 42]{protein}.
}
<<read_data>>=
yeast <- read.table("data/protein.txt")
yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
@
<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
@

>>>>

\maketitle
\begin{abstract}
Abstract will be here
\end{abstract}
<<chunk_setup, echo=FALSE>>=
library(knitr)
opts_chunk$set(tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
@
<<readingdata, eval=TRUE, echo=FALSE>>=
setwd("~/Dropbox/ggplot2/") 
library(ggplot2) # needs to be the version with geom_net!
 theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
 data(blood, football, theme_elements, lesmis)
names(blood$vertices)[1] <- 'label'
load("data/madmen.rda") # relative path should work for both of us
names(madmen$vertices)[1] <- 'label'
names(theme_elements$vertices)[1] <- 'label'
football$vertices$label <- as.factor(football$vertices$label)
football$vertices$value <- as.factor(football$vertices$value)
 email.net <- read.csv("data/email-data.csv")
 email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))

@@

This network is shown in figure~\ref{fig.cap:madmen_ex}.
<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net
@

@@

 In the plot, we can see one central character who has many more relationships than any other character. This vertex represents the main character of the show, Don Draper.
\par There are many kinds of networks, and networks are extensively studied across many disciplines.  Many sociologists study social networks, and many biologists study protein networks. As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.  We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}. \sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day. It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.}

@@

  \subsection{Vertex Aesthetics}
  In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.  The \texttt{vlabel} aesthetic is a logical value that prints the name or number assigned to each vertex in order to uniquely identify it on the plot. \sct{If this value is set to TRUE, the vertex data frame is required to have a column titled label.} This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be read clearly.  The \texttt{vcolor} aesthetic can be an identity color (e.g. \texttt{color = "red"}) to change the color of all vertices, \sct{XXX shouldn't we take this out? or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.}  \sct{Right now, the \texttt{vcolor} aesthetic does not work like a typical color aesthetic in \texttt{ggplot2}. It can, however, take HEX color characters and assign them to different levels of a factor variable.}  \sct{XXX We should probably take this out too. It can also color the vertices according to the different values of  some numeric variable, such as degree.  The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.}
  %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs. \sct{These must be integer-valued columns in the vertex data frame.} It can also be set to any other numeric variable associated with the vertices in the network.  The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.  It can also change the shape of the vertices according to some grouping variable, \sct{which is currently required to consist of integer values from 0 to 25.}.  Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10, or to some numeric variable of interest that takes on the values $[0,1]$.

\par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion.
  \subsection{Edge Aesthetics}
  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for lines.  \sct{XXX take out this??? The \texttt{elabel} aesthetic is the name or number assigned to each edge.  This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.}  The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to HEX colors assigned to levels of a grouping variable.  %It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge. 
  The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the different types of lines in \texttt{ggplot2}. \sct{The column associated with this aesthetic must be integer valued and take on values from one to six.}  The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function. \sct{The associated column can take on any numerical values.}  The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability. \sct{This property is especially useul for networks with a lot of connections and with vertices of relatively high degree.}  \sct{XXX should we take this out? Finally, the \texttt{directed} aesthetic is a logical value that identifies whether or not the network is directed.  When this value is true, the arrows change size, color, etc. with the previous aesthetics.}

@@

\section{Many Many Examples}
\sctq{Format-wise, is it ok if each example is its own subsection? Also, can you check the references I added to the .bib file to make sure I'm creating them correctly? I've just been appending them.}\\
\sct{
In this section, we demonstrate the capabilities of \texttt{geom\_net} in a series of examples.
  \subsection{Blood Donation}
  \emph{Note: would be best with labels and direction}\\
  In this directed network, there are eight vertices and 27 edges.  The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive and negative types. \sctq{Can I cite wikipedia here? Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a person whose blood type is that of a corresponding \emph{to} vertex. In the example below, loops are removed.  Loops exist on every vertex in this example, as blood of matching ABO and RhD type can always be received.
}
<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2), directed = TRUE) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add direction later
@
\sct{
This network is shown in figure \ref{fig.cap:blood_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
\end{figure}

\subsection{Email Network}
  \emph{Note: would be best with coloring and direction}
  This email network comes from the 2014 VAST Challenge \citep{emailnet}. It is a directed network of emails between company employees with 55 vertices and 9,063 edges. Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees. The arrow of the directed edge points to the recipient(s) of the email. The network contains two business weeks of emails across the entire company. }
<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
ggplot(data = email.net, aes(from_id = From, to_id = to)) +
  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  theme_net #add direction later
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
\end{figure}

\sct{This network is plotted in figure \ref{fig.cap:email_ex}.  This plot is very crowded, and it is difficult to distinguish structure in the network when it there is so much over-plotting.  To remedy this, we facet the network by day to examine how the company's communications change throughout the week. }
<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  facet_wrap(~day, nrow = 2) +
  theme_net #add direction later
@
\begin{landscape}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
\caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
\end{figure}
\end{landscape}
\sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.  \sctq{XXX note for later: remove all the emails that go to everyone in the network.} With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving? will know with direction option added later} several emails each day to a majority of the company. We can also see clusters that have formed on the perimeter of each day's network where the employees are communicating heavily with each other, but not much outside of their cluster. We can also see that there is heavier inter-cluster communication on different days.  For example, day 17 has more concentrated in-group communication than day 15. Additionally, we can better distinguish the communication of the outlying employee, who only corresponds with up to four other complany employees on a given day.} \sctq{talk about clusters being different departments within the company?}
  \subsection{\texttt{ggplot2} Theme Elements}
  \emph{Note: would be best with labels and direction} \\
  \sct{This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.  It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.  There are 53 vertices and 36 edges in this network.  There is an arrow from one vertex to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.  For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.  Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background. Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.  }
<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
  expand_limits(x = c(0,1.2), y = c(0,1)) +
  theme_net
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
\caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
\end{figure}

\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
  \subsection{Madmen Network (again?? or no?)}
  \sctq{leaving blank for now.}
  \subsection{College Football}
  \emph{Note: would be best with coloring vertices by conference}
  This next example comes from M.E.J. Newman's network data web page \citep{football}.  It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.  There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
ggplot(data = football$edges, aes(from_id = from, to_id = to, elinetype = in.conf + 1)) +
  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 2, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
  expand_limits(x = c(0,1), y = c(0,1)) +
  theme_net
@
\sct{
The network of football games is given in figure \ref{fig.cap:football_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
\end{figure}
  \subsection{Les Mis\'{e}rables}
  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
  This next network comes from \citet{lesmis}. It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.  There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.  There are 254 edges in this network. The edges are also weighted by the number of coappearances. The largest weighting is 31, between the characters Jean Valjean and Cosette.}
<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) + geom_net(vertices = lesmis$vertices, ealpha = 7/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
\end{figure}
\sct{
  \subsection{Protein Interaction Network in Yeast}
  This example of a protein interaction network comes from \citet{protein}. It is the complete protein-protein interaction network in the yeast species \emph{S. cerevisiae}. There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.  These edges represent ``direct physical interactions" between any 2 proteins \citep[][p. 42]{protein}.
}
<<read_data>>=
yeast <- read.table("data/protein.txt")
yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
@
<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
@
